---
title: "P8105 Homework 2"
output: github_document
---

```{r, echo = FALSE, message = FALSE}
library(tidyverse)
library(readxl)
```


Problem 1

The code chunk below reads and cleans Mr.Trash Wheel sheet.

```{r}
trash_wheel_sheet = read_excel("hw2_data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "Mr. Trash Wheel", range = "A2:N408") %>%
  janitor::clean_names() %>%
  drop_na(dumpster) %>%
  mutate(sports_balls = round(sports_balls))
```

The code chunk below reads and cleans precipitation data for 2018.
```{r}
precipitation_2018 = read_excel("hw2_data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2018 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  mutate(year = 2018)
```

The code chunk below reads and cleans precipitation data for 2019.
```{r}
precipitation_2019 = 
read_excel("hw2_data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = "2019 Precipitation", range = "A2:B14") %>%
  janitor::clean_names() %>%
  drop_na(total) %>%
  mutate(year = 2019)
```

The code chunk below combines precipitation datasets and converts month to a character variable.

```{r}
precipitation_combined = bind_rows(precipitation_2018, precipitation_2019) %>%
  mutate(month = month.name[month])
```

The number of observations in the resulting Mr.Trash Wheel dataset is `r nrow(trash_wheel_sheet)`. Names of key variables are `r names(trash_wheel_sheet)`. The number of observations in the resulting combined precipitation dataset is `r nrow(precipitation_combined)`. Names of key variables are `r names(precipitation_combined)`. The total precipitation in 2018 is `r sum(pull(precipitation_2018, total))`. The median number of sports balls in a dumpster in 2019 is `r median(pull(filter(trash_wheel_sheet, year == 2019), sports_balls))`.


Problem 2

The code chunk below deals with data in pols-month.csv.

```{r}
pols_month_data = read_csv("hw2_data/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon, into = c("year", "month", "day")) %>%
  mutate(year = as.integer(year), month = as.integer(month), day = as.integer(day)) %>%
  mutate(month = month.name[(month)]) %>%
  rename(dem = prez_dem) %>%
  rename(gop = prez_gop) %>%
  pivot_longer(c(dem, gop), names_to = "party", values_to = "president") %>%
  select(-day)
```

The code chunk below deals with data in snap.csv.

```{r}
snp_data = read_csv("hw2_data/fivethirtyeight_datasets/snp.csv") %>%
  separate(date, into = c("month", "day", "year")) %>%
  arrange(year, as.numeric(month))

snp_data_1 = filter(snp_data, year >= 50) %>%
  mutate(year = as.numeric(year) + 1900)

snp_data_2 = filter(snp_data, year <= 15) %>%
  mutate(year = as.numeric(year) + 2000)

snp_data_final = bind_rows(snp_data_1, snp_data_2) %>%
  mutate(month = month.name[as.integer(month)]) %>%
  select(-day) %>%
  relocate(year, month)
```

The code chunk below deals with unemployment data.

```{r}
unemployment_data = read_csv("hw2_data/fivethirtyeight_datasets/unemployment.csv") %>%
  pivot_longer(Jan:Dec, names_to = "month", values_to = "unemployment rate") %>%
  janitor::clean_names() %>%
  mutate(month = recode(month, `Jan` = "January", `Feb` = "February", `Mar` = "March", `Apr` = "April", `Jun` = "June", `Jul` = "July", `Aug` = "August", `Sep` = "September", `Oct` = "October", `Nov` = "November", `Dec` = "December"))
```

The code chunk below join the datasets by merging snp_data_final into pols_month_data, and merging unemployment_data into the result.

```{r}
polsmonth_snp = left_join(pols_month_data, snp_data_final)
polsmonth_snp_unemployment = left_join(polsmonth_snp, unemployment_data)
```

The pols-month dataset has dimension `r nrow(pols_month_data)` x `r ncol(pols_month_data)`. The range of years is from 1947 to 2015. Names of key variables are `r names(pols_month_data)`. The snp dataset has dimension `r nrow(snp_data_final)` x `r ncol(snp_data_final)`. The range of years is from 1950 to 2015. Names of key variables are `r names(snp_data_final)`. The unemployment dataset has dimension `r nrow(unemployment_data)` x `r ncol(unemployment_data)`. The range of years is from 1948 to 2015. Names of key variables are `r names(unemployment_data)`. The resulting dataset has dimension `r nrow(polsmonth_snp_unemployment)` x `r ncol(polsmonth_snp_unemployment)`. The range of years is from 1947 to 2015. Names of key variables are `r names(polsmonth_snp_unemployment)`.


Problem 3

The code chunk below loads and tidies popular baby names data.

```{r}
popular_baby_names = read_csv("hw2_data/Popular_Baby_Names.csv") %>%
  janitor::clean_names() %>%
  mutate(ethnicity = recode(ethnicity, `WHITE NON HISP` = "WHITE NON HISPANIC", `ASIAN AND PACI` = "ASIAN AND PACIFIC ISLANDER", `BLACK NON HISP` = "BLACK NON HISPANIC"), childs_first_name = str_to_title(childs_first_name))
```

The code chunk below removes duplicated rows in popular baby names dataset.

```{r}
popular_baby_names_filtered = popular_baby_names[!duplicated(popular_baby_names[,1:6]),]
```

The code chunk below produces a well-structured, reader-friendly table showing the rank in popularity of the name “Olivia” as a female baby name over time. 

```{r}
Olivia_female_popularity = filter(popular_baby_names_filtered, childs_first_name == "Olivia", gender == "FEMALE") %>%
  arrange(year_of_birth) %>%
  pivot_wider(ethnicity, names_from = "year_of_birth", values_from = "rank") %>%
  knitr::kable()
Olivia_female_popularity
```

The code chunk below produces a table showing the most popular name among male children over time.

```{r}
Most_popular_male_name = filter(popular_baby_names_filtered, rank == "1", gender == "MALE") %>%
  arrange(year_of_birth) %>%
  pivot_wider(ethnicity, names_from = "year_of_birth", values_from = "childs_first_name") %>%
  knitr::kable()
Most_popular_male_name
```